Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-14 PAPER.tex
made alpha into superscript instead of subscript in text right after eqn 2
changed caption to fig 6 and description to reflect new toy model

We are interested in estimating the three-dimensional power spectrum of the EoR.  
Visibilities are measurements of the Fourier transform of the sky along two spatial dimensions (using the flat-sky approximation), and since we are interested in three-dimensional Fourier modes we only need to take one Fourier transform of our visibilities along the line-of-sight dimension.  We consider band powers $p_\alpha$ of the power spectrum of $\textbf{x}$ over some range in cosmological $\mathbf{k}$, where $\alpha$ indexes a waveband in $k_{\parallel}$ (a cosmological wavenumber $k_{\parallel}$ is the Fourier dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}), which is a good approximation for the short baselines that PAPER analyzes).  The fundamental dependence of the covariance on the power spectrum band powers $p_\alpha$ is encoded as 
% JEA - Should distinguish slightly more clearly between Q and dC/dp_alpha
\begin{equation}
\textbf{S} = \sum_\alpha p^\alpha \frac{\partial\textbf{C}}{\partial p^\alpha} \equiv \sum_\alpha p^\alpha \textbf{Q}^\alpha
%\textbf{S} = p_{\alpha} \frac{\partial\textbf{C}}{\partial p_{\alpha}}
%\frac{\partial\textbf{C}}{\partial p_{\alpha}} \equiv \textbf{Q}^{\alpha}
\end{equation}
where we define $\frac{\partial\textbf{C}}{\partial p_\alpha} \equiv \textbf{Q}_{\alpha}$.
% JEA - ah, the dimensions again
% which has dimensions ($N_{f}$,$N_{f}$).

@@

%Just as important as the eigenvalues are the eigenvectors of our empirical covariances. 
The overall convergence of the covariance is important, but it is an important fact that, in general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix whose true form is a diagonal matrix with eigenvalues spanning four orders of magnitude. We use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$ for the empirical eigenvectors $\widehat{\textbf{v}}$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. One sees that the highest-valued eigenmodes converge to their true eigenvectors more quickly. With only a small number of realizations, these empirically estimated modes already retain little correlation with the specific realizations of data that were used to form the empirical covariance. As we will see in the next section, using only these eigenmodes, which are less coupled to data realizations, minimizes signal loss. In contrast, the lower-valued eigenmodes retain more memory of the data realizations, which leads to correlations that induce signal loss. 
% JEA - this is redundant
%Said differently, a steep covariance eigenspectrum can be especially dangerous because the eigenmodes that converge the slowest are typically also EoR-dominated and are therefore susceptible to the most loss.  %\acl{Think about whether we need to do the Fourier space version as well.}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss17.pdf}
	\caption{The convergence level, as defined in Equation \eqref{eq:converge_eig}, of empirically estimated eigenvectors for different numbers of mock data realizations. The colors span from the 0th eigenmode (has the highest eigenvalue) to the 19th eigenmode (has the lowest eigenvalue), where they are ordered by eigenvalue in descending order. This figure shows that the highest-valued modes converge more quickly than the others, implying that weighting by the lowest-valued empirically estimated modes poses the most risk for loss. Since these modes are also typically EoR-dominated, this can lead to cosmological signal loss.}
	\label{fig:toy_sigloss17}
\end{figure}

>>>>

We are interested in estimating the three-dimensional power spectrum of the EoR.  
Visibilities are measurements of the Fourier transform of the sky along two spatial dimensions (using the flat-sky approximation), and since we are interested in three-dimensional Fourier modes we only need to take one Fourier transform of our visibilities along the line-of-sight dimension.  We consider band powers $p_\alpha$ of the power spectrum of $\textbf{x}$ over some range in cosmological $\mathbf{k}$, where $\alpha$ indexes a waveband in $k_{\parallel}$ (a cosmological wavenumber $k_{\parallel}$ is the Fourier dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}), which is a good approximation for the short baselines that PAPER analyzes).  The fundamental dependence of the covariance on the power spectrum band powers $p_\alpha$ is encoded as 
% JEA - Should distinguish slightly more clearly between Q and dC/dp_alpha
\begin{equation}
\textbf{S} = \sum_\alpha p^\alpha \frac{\partial\textbf{C}}{\partial p^\alpha} \equiv \sum_\alpha p^\alpha \textbf{Q}^\alpha
%\textbf{S} = p_{\alpha} \frac{\partial\textbf{C}}{\partial p_{\alpha}}
%\frac{\partial\textbf{C}}{\partial p_{\alpha}} \equiv \textbf{Q}^{\alpha}
\end{equation}
where we define $\frac{\partial\textbf{C}}{\partial p^\alpha} \equiv \textbf{Q}^{\alpha}$.
% JEA - ah, the dimensions again
% which has dimensions ($N_{f}$,$N_{f}$).

@@

%Just as important as the eigenvalues are the eigenvectors of our empirical covariances. 
The overall convergence of the covariance is important, but it is an important fact that, in general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix. For this simulation, we construct a covariance whose ``true" form combines the same mock foreground from the previous toy models with an EoR component that is modeled as a diagonal matrix with eigenvalues spanning one order of magnitude (more specifically, we construct the EoR covariance as a diagonal matrix in the power spectrum domain, where the signal is expected to be uncorrelated; its Fourier-transform is then the ``true" covariance of the EoR in the visibility-domain, or $\textbf{C}_{\rm EoR}$). For different numbers of realizations, we draw random EoR signals that are consistent with $\textbf{C}_{\rm EoR}$, add them to the mock foreground data, and compute the combined empirical covariance (average over the realizations). The eigenvectors of this empirical covariance are then compared to the ``true" eigenvectors $\widehat{\textbf{v}}$, where we use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. We see that the zeroth eigenmode is quickest to converge --- this mode describes the foreground signal, which has a distinctly different covariance structure than that of the EoR. \cc{add discussion of eigenmode degeneracies and how that affects convergence rates... also state that the toy model happens to have a FG mode that dominates the highest eigenvalue, but it's not always that way}.

%One sees that the highest-valued eigenmodes converge to their true eigenvectors more quickly. With only a small number of realizations, these empirically estimated modes already retain little correlation with the specific realizations of data that were used to form the empirical covariance. As we will see in the next section, using only these eigenmodes, which are less coupled to data realizations, minimizes signal loss. In contrast, the lower-valued eigenmodes retain more memory of the data realizations, which leads to correlations that induce signal loss. 
% JEA - this is redundant
%Said differently, a steep covariance eigenspectrum can be especially dangerous because the eigenmodes that converge the slowest are typically also EoR-dominated and are therefore susceptible to the most loss.  %\acl{Think about whether we need to do the Fourier space version as well.}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/toy_sigloss17.pdf}
	\caption{The convergence level, as defined in Equation \eqref{eq:converge_eig}, of empirically estimated eigenvectors for different numbers of mock data realizations. The colors span from the 0th eigenmode (has the highest eigenvalue) to the 19th eigenmode (has the lowest eigenvalue), where they are ordered by eigenvalue in descending order. This figure shows that the 0th eigenmode converges the quickest, implying that eigenvectors that differ in structure from those of the EoR (i.e. the FG-dominated mode in the toy model has a different covariance structure than the EoR) converge to the true eigenvectors the quickest.}
	\label{fig:toy_sigloss17}
\end{figure}
